---
title: "Training Neural Networks with Evolutionary Algorithms"
author: "Benjamí Parellada and Clara Rivadulla"
date: "`r format(Sys.time(), '%d %b %Y')`"
output: 
  rmarkdown::html_document:
    toc: true
    number_sections: false
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(fig.align = "center", 
               out.width = "80%",
               fig.width = 6, fig.height = 5,
               dev = "svg", fig.ext = "svg",
               par = TRUE, # needed for setting hook 
               collapse = TRUE, # collapse input & output code in chunks
               warning = FALSE)

knit_hooks$set(par = function(before, options, envir)
  { if(before && options$fig.show != "none") 
       par(family = "sans", mar=c(4.1,4.1,1.1,1.1), mgp=c(3,1,0), tcl=-0.5)
})
```

## Introduction

## Dataset

As specified in the description of the exercise, we're going to use **synthetic data**. This will allow us to determine the sample size used for training, validation and testing, the true generalization error of a model, the amount of noise and the problem hardness.

We've decided to create a data set related to the performance of students in the Bachelor Degree in Informatics Engineering of the UPC. Specifically, we want to predict the grade that they will obtain in the subject of Programming Projects according to their previous results in two related subjects: Programming 1 and Programming 2.

First, we list down all variables we want to include in the data set, and describe the requirements of each one:


| Variable | Variable name | Description |
|-------|-------------|-------------|
| Student ID | id | Unique identifier of a student. 
| Programming 1 Grade* | prog1_grade | The grade that the student got in the Programming 1 subject. The average grade for all students is 6.4 out of 10.
| Programming 2 Grade* | prog2_grade | The grade that the student got in the Programming 2 subject. The average grade for all students is 6.7 out of 10.
| Class Absences | absences | The number of absences of a student (number of days they've not attended to class) during their first year. Students tend to have an average of 5 absences.
| Programming Projects Grade* | prog_proj_grade | The mean grade of a student from Semester 1 to Semester 4. The average mean grade is 7.3 out of 10. 
\* Variables that are correlated with one another

Now, we're going to define the functions that are useful for generating our synthetic data set: `generate_data`, to generate a dataframe of N data instances and split it into train, validation and test sets according to the fractions given (`fractionTraining`, `fractionValidation`, and `fractionTest`); and `round_df`, to round the data by defining a `max_value` and a `min_value`. 

In the generating process, we take into account that:
* To generate unique `IDs` is to create a sequence of whole numbers.
* The number of `absences` is distributed according to a Poisson distribution.
* The 3 grades (`prog1_grade`, `prog2_grade', `prog_proj_grade`) are correlated, and so we need a vector of means and a variance-covariance matrix to generate our multivariate normal distribution and calculate the covariances. 


```{r}
round_df <- function(df, digits, max_value, min_value) {
  nums <- vapply(df, is.numeric, FUN.VALUE = logical(1))

  df[,nums] <- round(df[,nums], digits = digits)
  df[,nums][df[,nums] > max_value] <- max_value
  df[,nums][df[,nums] < min_value] <- min_value

  (df)
}
```

```{r}
library(MASS)
generate_data <- function(N, fractionTraining, fractionValidation, fractionTest) {
  
  id <- paste0("S#", 1:N)
  set.seed(0717)
  absences <- rpois(N, lambda = 5)
  
  cor_var_means <- c(6.4, 6.7, 7.3)
  cor_var_matrix <- matrix(
    c(
      0.87, 0.65, 0.6,
      0.65, 1.2, 0.7,
      0.6, 0.7, 0.68
    ), byrow = T, nrow = 3
  )
  set.seed(0717)
  correlated_vars_df <- as.data.frame(mvrnorm(n = N, mu = cor_var_means, Sigma = cor_var_matrix))
  
  correlated_vars_df_cols <- c("prog1_grade", "prog2_grade", "prog_proj_grade")
  colnames(correlated_vars_df) <- correlated_vars_df_cols
  
  correlated_vars_df <- round_df(correlated_vars_df, digits = 1, max_value = 10, min_value = 0)
  
  df <- cbind(id, absences, correlated_vars_df)
  
  sampleSizeTraining   <- floor(fractionTraining   * nrow(df))
  sampleSizeValidation <- floor(fractionValidation * nrow(df))
  sampleSizeTest       <- floor(fractionTest       * nrow(df))

  indicesTraining    <- sort(sample(seq_len(nrow(df)), size=sampleSizeTraining))
  indicesNotTraining <- setdiff(seq_len(nrow(df)), indicesTraining)
  indicesValidation  <- sort(sample(indicesNotTraining, size=sampleSizeValidation))
  indicesTest        <- setdiff(indicesNotTraining, indicesValidation)


  dfTraining   <- df[indicesTraining, ]
  dfValidation <- df[indicesValidation, ]
  dfTest       <- df[indicesTest, ]
  
  return(list(dfTraining, dfValidation, dfTest))
  
}
```

We generate 3 different dataframes: one as the `train` set, one as the `validation` set and one as the `test` set. 

```{r}
data <- generate_data(3000, 0.8, 0.1, 0.1)
train <- data[[1]]
val <- data[[2]]
test <- data[[3]]
```


## Model

In order to define our **MLP** (*Multi Layer Perceptron*) model, we'll use the `nnet` package, which is intended for fitting feed-forward neural networks with a single hidden layer. 

```{r}
library(nnet)
model <- nnet(prog_proj_grade/10 ~ absences + prog1_grade + prog2_grade, data=train, size=5, decay=1.0e-5, maxit=2000)
model
summary(model)
```
```{r}
test_pred <- predict(model, test)*10
test_results <- data.frame(test_actual = test$prog2_grade, test_prediction = test_pred)
rmse <- sqrt(mean((test$prog2_grade-test_pred)^2))
rmse
```

Now, let's use GA to select the best hyperparameters of the Neural Network. We want to find the best value for the `size` of the `nnet` somewhere betweeen `1` and `2^5 = 32` and the best value for the `decay` parameter between `1.0e-1` and `1.0e-8`. To check how many bits we need, we multiply the maximum value of each hyperparameter and add it to the number of hyperparameters:

`log2(32 * 8) + 2 = 10`

```{r}
fit_rf=function(x)
{
  size=binary2decimal(x[1:6]) # size from 1 to 2^5
  decay=binary2decimal(x[7:10]) # decay from 1/10^1 to 1/10^8
  if(size==0)
  {
    size=1
  }
  if(decay==0)
  {
    decay=1
  }
  if(size>32)
  {
    size=32
  }
  if(decay>8)
  {
    decay=8
  }
  model=nnet(prog_proj_grade/10 ~ absences + prog1_grade + prog2_grade, data=train, size=size, decay=1.0/(10^decay))
  predictions=predict(model, test)
  rmse <- sqrt(mean((test$prog2_grade-predictions*10)^2))
  return(-rmse) #since GA maximize the objective function and we want to minimize RMSE
}
```

```{r}
library(GA)
library(tictoc)
tic()
GA=ga(type='binary',fitness=fit_rf,nBits=10,
       maxiter=30,popSize=50,seed=1234,keepBest=TRUE)
summary(GA)
plot(GA)
toc()
```
```{r}
ga_rf_fit=as.data.frame(GA@bestSol)
ga_size=apply(ga_rf_fit[, 1:6], 1, binary2decimal)
ga_size
ga_decay=apply(ga_rf_fit[, 7:10], 1, binary2decimal)
ga_decay
```

## Conclusions



## References
- https://towardsdatascience.com/how-to-create-a-custom-dataset-in-r-cf045e286656
- https://towardsdatascience.com/genetic-algorithm-in-r-hyperparameter-tuning-5fc6439d2962


